PREPRINT -- Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery -- PREPRINT.

### 2024-03-08 index.qmd
first part of Laurent's comments

  Coastal areas host seagrass meadows, which offer crucial ecosystem services including erosion control and carbon sequestration. However, these areas are increasingly impacted by human activities, leading to seagrass decline and habitat fragmentation. In situ surveys, traditionally performed to monitor these ecosystems face limitations on temporal and spatial coverage, particularly in intertidal zones, prompting the use of satellite data within monitoring programs. Yet, satellite remote sensing struggles with spatial and spectral resolution, making it difficult to discriminate seagrass from other macrophytes in highly heterogenous meadows. To address these challenges, drone images acquired with multi-spectral sensors offer a promising solution. This study focuses on using drones acquisitions for mapping intertidal macrophytes, effectively discriminating between seagrass and green macroalgae. Ten drone flights were conducted at two different altitudes (12m and 120m) across diverse European habitats in France and Portugal. Low altitude flights were used to train a Deep Learning classifier based on Neural Networks to discrimintate among 5 intertidal vegetation classes. Drone mapping demonstrated an overall accuracy of 94% across all the sites and images, covering a total area of 467 000 m². The model exhibited an accuracy of 96.4% in identifying seagrass.

@@

The importance of seagrass meadows and the variety of ecosystem services they provide have led to the enhancement of both global and regional programs monitoring Essential Oceanic Variable (EOVs) such as seagrass composition [@Miloslavich2018], as well as Essential Biodiversity Variable (EBVs) such as seagrass taxonomic diversity, species distribution, population abundance, and phenology [@Pereira2013]. Seagrass monitoring is all the more important as it makes possible to identify which pressure should be addressed in priority to facilitate effective mitigation actions. Traditionally, indicators of seagrass status have been quantified using in situ measurments. The acquisition of field data in intertidal zones is however notoriously challenging. Intertidal seagrass meadows are only partially exposed during low tide and can be situated in difficult-to-reach mudflats, potentially leading to inaccurate and limited estimations with conventional sampling techniques [@nijland2019]. Satellite observations have been proven effective in complementing in situ sampling, allowing for the near real-time and consistent retrieval of seagrass EOVs and EBVs over extensive meadows. [@Zoffoli2021 ; @xu2021 ; @Traganos2018 ; @coffer2023]

While satellite remote sensing provides temporally consistent observations over large spatial scales, it is limited by several constraints. Satellite missions with a high temporal resolution (e.g. daily MODIS observation) are limited by a coarse spatial resolution (\>100m). Mission with a high spatial resolution such as Sentinel-2 (10m) or Landsat8/9 (30m) can be limited by low spectral resolution. The limited number of spectral bands is indeed a challenge to accurately discriminate seagrass from others co-existing macrophytes. In particular Chlorophyceae (green algae) and marine Magnoliopsida (seagrass) share the same pigment composition [@ralph2002 ; @Douay2022]. As a result, their respective spectral signatures can be considered similar by a non-expert observer [@Davies2023 ; @bannari2022]. Recently, using advanced machine-learning algorithms trained with a large hyperspectral library of more then 300 field reflectance measurments, @Davies2023 demonstrated that it was possible to discriminate Magnoliopsida from Chlorophyceae using radiometric data acquired at Sentinel-2 's spectral resolution. The application of this approach to satellite remote sensing remains however to be validated. Moreover patches of green algae can develop at small spatial scales that are not observable using non-commercial satellite imagery [@tuya2013], especially during the initial stage of the event.

Drones can potentially fill the gaps left by satellite remote sensing and in situ measurments, due to their ability to provide spatially-explicit observations at very high spatial resolutions (pixel size from cm to mm) and still capturing data at multi-spectral resolution [@fairley2022drone ; @oh2017use]. The versatility of drones allows for their application across a diverse thematic range , from coastal zone management [@adade2021 ; @casella2020 ; @angnuureng2022] to mapping species distribution [@joyce2023 ; @tallam2023 ; @Roca2022 ; @Roman2021 ; @Brunier2022Topographic]. However, when applied to coastal habitat mapping, most study cases were limited to a single flight, restricting the generalizability of their application over wider geographical scales [@Roman2021 ; @collin2019improving ; @rossiter2020uav ; @Brunier2022Topographic]. The present study aimed at analyzing the potential of multispectral drone remote sensing to map intertidal macrophytes, with a particular focus on discriminating Magnoliopsida and Chlorophyceae. Ten drone flights were performed over soft-bottom intertidal areas along two European countries (France and Portugal), covering a wide range of habitats, from monospecific seagrass meadows to meadows mixed with green or red algae. A deep learning algorithm was trained and validated for macrophyte discrimination, emphasizing applicability across diverse sites without a loss of prediction accuracy.

@@

Seven study sites distributed between France and Portugal were selected for their relatively extensive intertidal seagrass beds. Two sites are located in Gulf of Morbihan (@fig-map A), France (47.5791°N, 2.8018°W). This gulf covers an area of 115 km² and is only connected to the sea through a 900m wide channel. A total of 53 small islands are scattered across the gulf leading to 250 km of shorelines. Patchy seagrass meadows can be found on a lot of these islands. Within the Gulf of Morbihan we have chosen two sites : one is located in one of these island (Arz island); the other one is located in the southern part of the gulf (Duer). Two others sites are located in Bourgneuf Bay, France (46.9849°N, 2.1488°W). This bay is a semi-enclosed macrotidal bay, protected from waves by Noirmoutier Island. Bougneuf bay hosts a large intertidal seagrass meadow of about 6 km². Within this meadow, the sites observed by drones (L'Epine and Barbatre, @fig-map B) contain a monospecific bed of *Zostera noltei* with very little mixing with other macrophytes. Three sites have been surveyed in the Ria de Aveiro Lagoon in Portugal (40.6887°N, 8.6810°W). The extent of this lagoon is about 80 km² with many narrow channels, large salt marshes and many mudflats that uncovers at low tide. It is connected to the open sea through a single channel, with a tidal lag between the North and the South of the lagoon. The southernmost site (Gafanha) is a mudflat located in an affluent of the lagoon whereas the two other sites are situated in the middle of the lagoon and are only accessible by boat (@fig-map C). These Portuguese sites are characterized by a more diverse intertidal vegetation, where patches of seagrass could intermingle with red algae (Rhodophyceae), brown algae (Phaeophyceae), and green algae (Chlorophyceae).

@@

```{r Figure1}
#| cache: false
#| echo: false
#| warning: false
#| fig-cap: >
#|   Sites observed by drone. A: Gulf of Morbihan, B: Bourngeuf Bay, C: Ria de Aveiro Lagoon. Light green represent terrestrial areas whereas darkgreen represent intertidal areas. 
#| label: fig-map
#| out-width: "100%"

@@

At each location, a DJI Matrice 200 quadcopter drone equipped with a Micasense RedEdge Dual MX multispectral camera was flown to take 1.2 million pixel reflectance photographs in 10 spectral bands, from 444 nm (blue) to 840 nm (near infrared, NIR). An angle of 90° was maintained between the sun and the drone's route to guarantee uniform lighting conditions between flight lines. A side overlap of 70% and a front overlap of 80% between each image was set for each flight. A downwelling light sensor (DLS2) was used to acquire irradiance data concomitantly with the camera measurements. Raw data were calibrated in reflectance using a calibration panel reflective at \~50% provided by the manufacturer. A structure-from-motion photogrammetry software (Agisoft Metashape) was used to process images to obtain multispectral orthomosaics of each flight. The workflow for orthomosaicking was the same for every flight. First, tying key points were detected inside of each image and between overlapping images in order to obtain a sparse point cloud. This cloud was cleaned using reprojection accuracy metric in order to remove noisy points. A dense point cloud was then produced using a structure from motion algorithm. A surface interpolation of this dense point cloud was made to obtain a digital surface model (DSM), used to reconstruct the multispectral ortho-image. Across all sites, flights were made at two different altitudes : 12 m or/and 120 m. Low altitude drone flights produce ortho-images with a very high spatial resolution (8 mm per pixel), making it simple to visually distinguish between the various types of vegetation. High altitude fights on the other hand allow to cover large areas and produced images with a pixel size of 80 mm (Table 1).

```{r Table1}
#| echo: false
#| warning: false
#| tbl-cap: List of drone flight, summarising the date, the altitude and the purpose of each flight.
#| cache: true

df<-data.frame(
  Country = c(
    rep("France",5),
    rep("Portugal",4)
  ),
  Site = c(
    rep("Morbihan",3),
    rep("Bourgneuf Bay",2),
    rep("Aveiro Lagoon",4)
  ),
  Name = c(
    "Arz Island",
    "Duer",
    "Duer",
    "Barbâtre",
    "L'Epine",
    "Inner Lagoon",
    "Mataducos",
    "Gafanha",
    "Gafanha"
  ),
  Altitude = c(
    "12m",
    "12m",
    "120m",
    "120m",
    "120m",
    "120m",
    "120m",
    "120m",
    "12m"
  ),
  Utility = c(
    "Training",
    "Training",
    "Validation",
    "Validation",
    "Validation",
    "Validation",
    "Validation",
    "Validation",
    "Training"
  )
)

@@

Before each flight, targets used as Ground Control Points (GCPs) were distributed over the study site and georeferenced with a Trimble © Geo XH 6000 differential GPS (dGPS). GCPs were used to correct georeferencing imprecision of orthomosaics with an horizontal and vertical accuracy of 10cm. dGPS was also used to georeference quadrats of 0.25 m² used to assess the presence or absence of 5 key classes of vegetation : Bacillariophyceae (Benthic diatoms forming large biofilm over the sediment during low tide), Phaeophyceae (brown algae), Magnoliopsida (seagrass), Chlorophyceae (green algae) and Rhodophyceae (red algae) (@fig-vegetation). Pictures of each quadrat were uploaded online on the Global Biodiversity Information Facility (GBIF) platform [@BedeGbif], a public open portal to store and share biodiversity data. Each photograph was also processed to estimate the percent cover of each type of vegetation using an image processing software (imageJ). For all quadrat, the hyperspectral reflectance signatures of each vegetation class was measured using an ASD FieldSpec HandHeld 2 spectroradiometer, which acquires reflectance between 325 and 1075 nm, with 1 nm of spectral resolution. The methodological scheme followed in this study is presented in @fig-workflow.

@@

```{r Figure2 Plot}
#| cache: true
#| echo: false
#| warning: false
#| fig-cap: Class of vegetation used to train the Neural Network model and their standardised spectral signature
#| label: fig-vegetation
#| out-width: "100%"

@@

```{r Figure3}
#| cache: false
#| echo: false
#| warning: false
#| fig-cap: "Schematic representation of the workflow. Diamonds represent input or output data, and rectangles represent Python processing algorithms. The overall workflow of this study is divided into two distinct parts based on the spatial resolution of the drone flights: high-resolution flights were utilized for training and prediction of the Neural Network model, whereas low-resolution flights were solely employed for prediction purposes."
#| label: fig-workflow
#| fig-width: 10
#| fig-height: 7
#| out-width: "100%"

@@

df_validation<-"Data/accuracy_assesment_low&high_flights.csv" %>% 
  read_delim(delim = ",", escape_double = FALSE, trim_ws = TRUE) %>% 
  dplyr::filter(True != "Sediment",
                True != "Sun Glint",
                True != "Water",
                True != "Xanthophyceae") %>% 
  mutate(True = case_when(True == "Green Algae" ~ "Chlorophyceae",
                          True == "Microphytobenthos" ~ "Bacillariophyceae",
                          True == "Rhodophyceae" ~ "Rhodophyceae",
                          True == "Pheophyceae" ~ "Phaeophyceae",
                          TRUE ~ True)) %>% 
  group_by(True) %>% 
  summarise(n = n())

@@

```{r Table2}
#| echo: false
#| warning: false
#| tbl-cap: Vegetation Classes of the model and the number of pixels used to train and validate each class
#| cache: true

@@

A dataset containing photo-interpreted drone reflectance pixels was built to train a Neural Network model with 2 hidden layers. The training pixels were categorized into 7 different classes, representing the various habitats encountered at the different study sites: Sediment, Water, Chlorophyceae, Magnoliopsida, Bacillariophyceae, Phaeophyceae and Rhodophyceae. Only low-altitude flights (Table 1) were used for training purposes because their 8 mm spatial resolution allowied to avoid spectral sub-pixel mixing and to accurately differentiate various vegetation classes. More than `r my_comma(round(sum(df_training$Training_Pixels),-3))` pixels at 8 mm resolution from the 3 training flights were used to train the model (Table 2). Twenty one variables were used by the model as predictors: 10 raw spectral bands of the Micasense RedEdge Dual MX multispectral camera (ranging from 444 nm to 840 nm), the same 10 spectral bands standardized using a min/max transformation ([@eq-std] ; @Cao2017) and the Normalized difference vegetation index (NDVI, @eq-ndvi). Standardisation of spectral bands is used to eliminate the scaling differences between spectra and to limit the effect of biomass on the shape of the spectra [@Douay2022 ; @Davies2023].

@@

>>>>

  Coastal areas host seagrass meadows, which offer crucial ecosystem services including erosion control and carbon sequestration. However, these areas are increasingly impacted by human activities, leading to seagrass decline and habitat fragmentation. In situ surveys, traditionally performed to monitor these ecosystems face limitations on temporal and spatial coverage, particularly in intertidal zones, prompting the use of satellite data within monitoring programs. Yet, satellite remote sensing struggles with spatial and spectral resolution, making it difficult to discriminate seagrass from other macrophytes in highly heterogenous meadows. To address these challenges and reduce the issue of mlixed pixel, drone images at a very high spatial resolution offer a promising solution. This study focuses on using drone acquisitions with a ten spectral bands sensor mirroring those of Sentinel-2, for mapping intertidal macrophytes and effectively discriminating between seagrass and green macroalgae. Nine drone flights were conducted at two different altitudes (12m and 120m) across diverse European habitats in France and Portugal. Low altitude flights were used to train a Deep Learning classifier based on Neural Networks to discrimintate among five taxonomic classes of intertidal vegetation: Magnoliopsida (Seagrass), Chlorophyceae (Green macroalgae), Phaeophyceae (Brown algae), Rhodophyceae (Red macroalgae) and Bacillariophyceae (Diatoms). Drone mapping demonstrated an overall accuracy of 94% across all the sites and images, covering a total area of 467 000 m². The model exhibited an accuracy of 96.4% in identifying seagrass. This work showed that although they share the same pigmentary composition, seagrass and green algae can be discriminated using ten bands multispectral sensor. The algorithm should now be adapted for Sentinel-2 to see if this discrimination is still operational at a coarser spatial resolution. 

@@

The importance of seagrass meadows and the variety of ecosystem services they provide have led to the enhancement of both global and regional programs monitoring Essential Oceanic Variable (EOVs) such as seagrass composition [@Miloslavich2018], as well as Essential Biodiversity Variable (EBVs) such as seagrass taxonomic diversity, species distribution, population abundance, and phenology [@Pereira2013]. Seagrass monitoring is all the more important as it makes possible to identify which pressure should be addressed in priority to facilitate effective mitigation actions. Traditionally, indicators of seagrass status have been quantified using *in situ* measurements. The acquisition of field data in intertidal zones is however notoriously challenging. Intertidal seagrass meadows are only partially exposed during low tide and can be situated in difficult-to-reach mudflats, potentially leading to inaccurate and limited estimations with conventional sampling techniques [@nijland2019]. Satellite observations have been proven effective in complementing in situ sampling, allowing for the near real-time and consistent retrieval of seagrass EOVs and EBVs over extensive meadows. [@Zoffoli2021 ; @xu2021 ; @Traganos2018 ; @coffer2023]

While satellite remote sensing provides temporally consistent observations over large spatial scales, it is limited by several constraints. Satellite missions with a high temporal resolution (e.g. daily MODIS observation) are limited by a coarse spatial resolution (\>100m). Mission with a high spatial resolution such as Sentinel-2 (10m) or Landsat8/9 (30m) can be limited by low spectral resolution. The limited number of spectral bands is indeed a challenge to accurately discriminate seagrass from others co-existing macrophytes. In particular Chlorophyceae (green algae) and marine Magnoliopsida (seagrass) share the same pigment composition [@ralph2002 ; @Douay2022]. As a result, their respective spectral signatures can be considered similar by a non-expert observer [@Davies2023 ; @bannari2022]. Recently, using advanced machine-learning algorithms trained with a large hyperspectral library of more then 300 field reflectance spectra, @Davies2023 demonstrated that it was possible to discriminate Magnoliopsida from Chlorophyceae using radiometric data acquired at Sentinel-2 's spectral resolution. The application of this approach to satellite remote sensing remains however to be validated. Moreover patches of green algae can develop at small spatial scales that are not observable using non-commercial satellite imagery [@tuya2013], especially during the initial stage of the event.

Drones can potentially fill the gaps left by satellite remote sensing and in situ measurements, due to their ability to provide spatially-explicit observations at very high spatial resolutions (pixel size from cm to mm) and still capturing data at multi-spectral resolution [@fairley2022drone ; @oh2017use]. The versatility of drones allows for their application across a diverse thematic range , from coastal zone management [@adade2021 ; @casella2020 ; @angnuureng2022] to mapping species distribution [@joyce2023 ; @tallam2023 ; @Roca2022 ; @Roman2021 ; @Brunier2022Topographic]. However, when applied to coastal habitat mapping, most study cases were limited to a single flight, restricting the generalizability of their application over wider geographical scales [@Roman2021 ; @collin2019improving ; @rossiter2020uav ; @Brunier2022Topographic]. The present study aimed at analyzing the potential of multispectral drone remote sensing to map intertidal macrophytes, with a particular focus on discriminating Magnoliopsida and Chlorophyceae. Ten drone flights were performed over soft-bottom intertidal areas along two European countries (France and Portugal), covering a wide range of habitats, from monospecific seagrass meadows to meadows mixed with green, or red macroalgae. A deep learning algorithm was trained and validated for macrophyte discrimination, emphasizing applicability across diverse sites without a loss of prediction accuracy.

@@

Seven study sites distributed between France and Portugal were selected for their relatively extensive intertidal seagrass beds. Two sites are located in Gulf of Morbihan (@fig-map A), France (47.5791°N, 2.8018°W). This gulf covers an area of 115 km² and is only connected to the sea through a 900m wide channel. A total of 53 small islands are scattered across the gulf leading to 250 km of shorelines. Patchy seagrass meadows can be found on a lot of these islands. Within the Gulf of Morbihan we have chosen two sites : one is located in one of these island (Arz island); the other one is located in the southern part of the gulf (Duer). Two others sites are located in Bourgneuf Bay, France (46.9849°N, 2.1488°W). This bay is 340 km² semi-enclosed macrotidal bay, protected from waves by Noirmoutier Island. Bougneuf bay hosts a large intertidal seagrass meadow of about 6 km². Within this meadow, the sites observed by drones (L'Epine and Barbatre, @fig-map B) contain a monospecific bed of *Zostera noltei* with very little mixing with other macrophytes. Three sites have been surveyed in the Ria de Aveiro Lagoon in Portugal (40.6887°N, 8.6810°W). The extent of this lagoon is about 80 km² with many narrow channels, large salt marshes and many mudflats that uncovers at low tide. It is connected to the open sea through a single channel, with a tidal lag between the North and the South of the lagoon. The southernmost site (Gafanha) is a mudflat located in an affluent of the lagoon whereas the two other sites are situated in the middle of the lagoon and are only accessible by boat (@fig-map C). These Portuguese sites are characterized by a more diverse intertidal vegetation, where patches of seagrass could intermingle with red, brown and green macroalgae.

@@

```{r Figure1}
#| cache: false
#| echo: false
#| warning: false
#| fig-cap: >
#|   Location of drone flights in France and Portugal. A: Gulf of Morbihan (Two sites), B: Bourngeuf Bay (Two sites), C: Ria de Aveiro Lagoon (Three sites). Green represent intertidal areas. 
#| label: fig-map
#| out-width: "100%"

@@

At each location, a DJI Matrice 200 quadcopter drone equipped with a Micasense RedEdge Dual MX multispectral camera was flown to take 1.2 million pixel reflectance photographs in ten spectral bands, from 444 nm (blue) to 840 nm (near infrared, NIR). An angle of 90° was maintained between the sun and the drone's route to guarantee uniform lighting conditions between flight lines. A side overlap of 70% and a front overlap of 80% between each image was set for each flight. A downwelling light sensor (DLS2) was used to acquire irradiance data concomitantly with the camera measurements. Raw data were calibrated in reflectance using a calibration panel reflective at \~50% provided by the manufacturer. A structure-from-motion photogrammetry software (Agisoft Metashape) was used to process images to obtain multispectral orthomosaics of each flight. The workflow for orthomosaicking was the same for every flight. First, tying key points were detected inside of each image and between overlapping images in order to obtain a sparse point cloud. This cloud was cleaned using reprojection accuracy metric in order to remove noisy points. A dense point cloud was then produced using a structure from motion algorithm. A surface interpolation of this dense point cloud was made to obtain a digital surface model (DSM), used to reconstruct the multispectral ortho-image. Across all sites, flights were made at two different altitudes : 12 m or/and 120 m. Low altitude drone flights produce ortho-images with a very high spatial resolution (8 mm per pixel), making it efficient to visually distinguish between the various types of vegetation. High altitude fights on the other hand allow to cover large areas and produced images with a pixel size of 80 mm (@tbl-flights).

```{r Table1}
#| echo: false
#| warning: false
#| label: tbl-flights
#| tbl-cap: List of drone flight, summarising the date, the altitude and the purpose of each flight. 12 m  and 120 m flights have a spatial resolution of 8 and 80 mm respectively.
#| cache: true

df<-data.frame(
  Country = c(
    rep("France",5),
    rep("Portugal",4)
  ),
  Site = c(
    rep("Gulf of Morbihan",3),
    rep("Bourgneuf Bay",2),
    rep("Aveiro Lagoon",4)
  ),
  Name = c(
    "Arz Island",
    "Duer",
    "Duer",
    "Barbâtre",
    "L'Epine",
    "Inner Lagoon",
    "Mataducos",
    "Gafanha",
    "Gafanha"
  ),
  Altitude = c(
    "12m",
    "12m",
    "120m",
    "120m",
    "120m",
    "120m",
    "120m",
    "120m",
    "12m"
  ),
  Utility = c(
    "Training",
    "Training",
    "Validation",
    "Validation",
    "Validation",
    "Validation",
    "Validation",
    "Validation",
    "Training"
  )
)

@@

Before each flight, targets used as Ground Control Points (GCPs) were distributed over the study site and georeferenced with a Trimble © Geo XH 6000 differential GPS (dGPS). GCPs were used to correct georeferencing imprecision of orthomosaics with an horizontal and vertical accuracy of 10cm. dGPS was also used to georeference quadrats of 0.25 m² assessing the presence or absence of five key taxonomic classes of intertidal vegetation : Bacillariophyceae (Unicellular benthic diatoms forming large biofilms at the sediment surface during low tide), Phaeophyceae (brown macroalgae), Magnoliopsida (seagrass), Chlorophyceae (green macroalgae) and Rhodophyceae (red macroalgae) (@fig-vegetation). Pictures of each quadrat were uploaded online on the Global Biodiversity Information Facility (GBIF) platform [@BedeGbif], a public open portal to store and share biodiversity data. Each photograph was also processed to estimate the percent cover of each type of vegetation using an image processing software (imageJ). For all quadrats, the hyperspectral reflectance signatures of each vegetation class was measured using an ASD FieldSpec HandHeld 2 spectroradiometer, which acquires reflectance between 325 and 1075 nm, with 1 nm of spectral resolution. Hyperspectral signatures from quadrats serve dual purposes: they validate the radiometric calibration of drone data and contribute to error reduction in photo interpretation. The workflow developped in this study is presented in @fig-workflow.

@@

```{r Figure2 Plot}
#| cache: true
#| echo: false
#| warning: false
#| fig-cap: The five taxonomic classes of vegetation used to train the Neural Network model and their raw spectral signatures at the spectral resolution of the Micasense RedEdge Dual MX. 
#| label: fig-vegetation
#| out-width: "100%"

@@

```{r Figure3}
#| cache: false
#| echo: false
#| warning: false
#| fig-cap: "Schematic representation of the workflow. Diamonds represent input or output data, and rectangles represent Python processing algorithms. The overall workflow of this study is divided into two distinct parts based on the spatial resolution of the drone flights: high-resolution flights (pixel size: 8 mm) were utilized for training and prediction of the Neural Network model, whereas lower-resolution flights (pixel size: 80 mm) were solely employed for prediction purposes. Validation has been performed on both high and low resolution flights."
#| label: fig-workflow
#| fig-width: 10
#| fig-height: 7
#| out-width: "100%"

@@

df_validation<-"Data/accuracy_assesment_low&high_flights.csv" %>% 
  read_delim(delim = ",", escape_double = FALSE, trim_ws = TRUE) %>% 
  dplyr::filter(True != "Sediment",
                True != "Sun Glint",
                True != "Water",
                True != "Xanthophyceae") %>% 
  mutate(True = case_when(True == "Green Algae" ~ "Chlorophyceae",
                          True == "Microphytobenthos" ~ "Bacillariophyceae",
                          True == "Rodophyceae" ~ "Rhodophyceae",
                          True == "Pheophyceae" ~ "Phaeophyceae",
                          TRUE ~ True)) %>% 
  group_by(True) %>% 
  summarise(n = n())

@@

```{r Table2}
#| echo: false
#| warning: false
#| tbl-cap: Vegetation Classes of the model and the number of pixels used to train and validate each class
#| label: tbl-validationPX
#| cache: true

@@

A dataset containing photo-interpreted drone reflectance pixels was built to train a Neural Network model with 2 hidden layers. The training pixels were categorized into seven different classes, representing the various habitats encountered at the different study sites: Sediment, Water, Chlorophyceae, Magnoliopsida, Bacillariophyceae, Phaeophyceae and Rhodophyceae. Only low-altitude flights (@tbl-flights) were used for training because their 8 mm spatial resolution allowied to avoid spectral sub-pixel mixing and to accurately differentiate vegetation classes. More than `r my_comma(round(sum(df_training$Training_Pixels),-3))` pixels at 8 mm resolution from the 3 training flights were used to train the model (@tbl-validationPX). Twenty one variables were used by the model as predictors: the ten raw spectral bands of the Micasense RedEdge Dual MX multispectral camera (ranging from 444 nm to 840 nm), the same ten spectral bands standardized using a min/max transformation ([@eq-std] ; @Cao2017) and the Normalized difference vegetation index (NDVI, @eq-ndvi). Standardisation of spectral bands is used to eliminate the scaling differences between spectra and to limit the effect of biomass on the shape of the spectra [@Douay2022 ; @Davies2023].

@@

The workflow of this study revolves around two distinct flight heights (12 m and 120 m, @fig-workflow) where ensuring consistency between reflectances at both heights is crucial. This comparison was conducted at sites where low and high altitude flights overlapped. The low altitude flights were resampled to the same spatial resolution and grid as the high flights using a median resampling method. Reflectance values were then extracted, and a scatterplot was generated. Subsequently, the slope of the linear model and the coefficient of determination (R²) were computed.
